---
title: "The R4SUB Evidence Schema"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The R4SUB Evidence Schema}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Every package in the R4SUB ecosystem reads and writes the same evidence table.
A quality tool, a traceability engine, and a risk model all emit rows in this
one shape, which is what lets `r4subscore` combine them into a single readiness
score. This article is the formal specification of that table.

```{r load}
library(r4subcore)
```

## The columns

`evidence_schema_spec()` returns the schema as a table: one row per column, its
type, whether it is required, its controlled vocabulary if it has one, and a
short description.

```{r spec}
spec <- evidence_schema_spec()
knitr::kable(spec)
```

A row is one observation: a single check, run against a single asset, producing
a single result. Required columns must be present and non-missing. Nullable
columns may be `NA`.

## Controlled vocabularies

Four columns are restricted to a fixed set of values. Using the canonical set
keeps scoring and aggregation consistent across sources.

```{r vocab}
schema <- evidence_schema()
schema$asset_type$allowed
schema$indicator_domain$allowed
schema$severity$allowed
schema$result$allowed
```

`severity` runs from `info` (no weight) to `critical` (full weight).
`indicator_domain` names the readiness pillar the check feeds: quality, trace,
risk, or usability. `result` is the outcome, where `na` means the check did not
apply.

If your source uses different labels, `canon_severity()` and `canon_result()`
map common aliases onto the canonical set:

```{r canon}
canon_severity(c("HIGH", "warning", "Error"))
canon_result(c("PASS", "Failed", "N/A"))
```

## Run context

`run_id` and `study_id` tie every row back to the run that produced it. Rather
than set them by hand on each row, create a run context once and let
`as_evidence()` fill them in:

```{r context}
ctx <- r4sub_run_context(study_id = "STUDY01", environment = "DEV")

ev <- as_evidence(
  data.frame(
    asset_type = "validation",
    asset_id = "ADSL",
    source_name = "pinnacle21",
    indicator_id = "SD0001",
    indicator_name = "Missing variable",
    indicator_domain = "quality",
    severity = "high",
    result = "fail",
    message = "Variable AGEU is missing",
    stringsAsFactors = FALSE
  ),
  ctx = ctx
)

ev[, c("run_id", "study_id", "indicator_id", "severity", "result")]
```

## Validation

`as_evidence()` validates as it builds, and `validate_evidence()` checks an
existing table against the specification: required columns present, types
correct, and controlled vocabularies respected. A table that passes
`validate_evidence()` is guaranteed to score.

```{r validate}
validate_evidence(ev)
```

Because the specification is a function, downstream packages can rely on it
directly rather than restating the column list. That is what keeps the ecosystem
in step: there is one schema, and this is it.
