---
title: "Reading Pinnacle 21 reports into evidence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reading Pinnacle 21 reports into evidence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Most clinical programming teams already run Pinnacle 21 on every submission and
have the report on disk. That report is the fastest way to get real evidence
into R4SUB: there is no table to build by hand. This vignette shows the two
steps, reading the report and converting it to the shared evidence table.

```{r load}
library(r4subcore)
```

## A Pinnacle 21 export

A Pinnacle 21 report lists one row per finding: a rule identifier, a message, a
severity, the dataset and variable it concerns, and a status. Here we write a
small export to a temporary CSV so the vignette is self-contained. Your real
CSV or Excel export has the same shape.

```{r export}
p21 <- data.frame(
  "Rule ID" = c("SD0001", "SD0002", "AD0018", "SD1077"),
  Message   = c("Variable label is missing",
                "Value not in controlled terminology",
                "Expected variable is present",
                "Inconsistent value length"),
  Severity  = c("Error", "Warning", "Notice", "Error"),
  Dataset   = c("ADSL", "ADAE", "ADSL", "ADLB"),
  Variable  = c("AGEU", "AESEV", "RACE", "LBORRES"),
  Status    = c("Failed", "Warning", "Passed", "Failed"),
  check.names = FALSE
)

csv <- tempfile(fileext = ".csv")
utils::write.csv(p21, csv, row.names = FALSE)
```

## Reading the report

`read_p21_report()` loads the file. It reads CSV with base R and Excel through
the `readxl` package when that is installed. It only reads the file; it does not
interpret the columns yet.

```{r read}
report <- read_p21_report(csv)
report
```

## Converting to evidence

`p21_to_evidence()` maps the report onto the evidence schema shared across the
ecosystem. It detects the rule, message, severity, dataset, variable, and result
columns case-insensitively, so it tolerates the small naming differences between
Pinnacle 21 versions. Pinnacle 21 severities map onto the evidence vocabulary:
`Error` becomes `high`, `Warning` becomes `medium`, and `Notice` becomes `info`.

```{r evidence}
ctx <- r4sub_run_context("STUDY1", "DEV")
ev  <- p21_to_evidence(report, ctx)

ev[, c("indicator_id", "severity", "result", "location")]
```

The result is an ordinary evidence table. It validates against the schema, so it
scores like any other evidence, and it can be combined with evidence from other
sources (Define-XML, the ADRG, traceability) with `bind_evidence()` before
computing a Submission Confidence Index.

```{r validate}
validate_evidence(ev)
```
